### 2012-12-10 inst/examples/mcmc-gp.Rmd
trouble-shooting priors

@@

I might try something like

```{r}
library(tgp)
gp <- bgp(X=x, XX=X, Z=y, verb=0,
          meanfn="constant", bprior="bflat", BTE=c(1,2,1), m0r1=FALSE, 
          corr="exp", trace=TRUE, beta = 0,
          s2.p = c(25,10), d.p = c(50, 10, 10, 10), nug.p = c(50, 10, 10, 10),
          s2.lam = "fixed", d.lam = "fixed", nug.lam = "fixed", 
          tau2.lam = "fixed", tau2.p=c(5,10))
```

@@

```{r}
V <- gp$ZZ.ks2
tgp_dat <- data.frame(x   = gp$XX[[1]], 
                  y   = gp$ZZ.km, 
                 ymin = gp$ZZ.km - 1.96 * sqrt(gp$ZZ.ks2), 
                 ymax = gp$ZZ.km + 1.96 * sqrt(gp$ZZ.ks2))

@@

```{r gp-plot}
library(ggplot2)
ggplot(tgp_dat) +
    geom_ribbon(aes(x, y, ymin = ymin, ymax = ymax), fill="red", alpha = .1) + # Var
    geom_line(aes(x, y), col="red") + # mean
    geom_point(data = data.frame(x = x, y = y), aes(x, y)) + # raw data
    geom_ribbon(dat = manual_dat, aes(x, y, ymin = ymin, ymax = ymax), fill = "blue", alpha = .1) + # Var
    geom_line(dat = manual_dat, aes(x, y), col = "blue")  + #MEAN    
    theme_bw() + theme(plot.background = element_rect(fill = "transparent",colour = NA))
```

>>>>

```{r libraries}
library(MCMCpack)
library(tgp)
library(reshape2)
library(ggplot2)
```

@@

## Priors

```{r}
s2.p <- c(50,50)
tau2.p <- c(20,1)
d.p = c(10, 1/0.01, 10, 1/0.01)
nug.p = c(10, 1/0.01, 10, 1/0.01)
```

Define as random samples (avoids having to specify range)

```{r}
s2_rprior <- function(x) rinvgamma(x, s2.p[1], s2.p[2])
tau2_rprior <- function(x) rinvgamma(x, tau2.p[1], tau2.p[2])
d_rprior <- function(x) rgamma(x, d.p[1], scale = d.p[2]) + rgamma(x, d.p[3], scale = d.p[4])
nug_rprior <- function(x) rgamma(x, nug.p[1], scale = nug.p[2]) + rgamma(x, nug.p[3], scale = nug.p[4])
beta0_rprior <- function(x, tau) rnorm(x, 0, tau)
n = 1000
rpriors <- data.frame(s2 = s2_rprior(n), tau2 = tau2_rprior(n), beta0 = beta0_rprior(n, 1), nug = nug_rprior(n), d = d_rprior(n))
rpriors <- melt(rpriors)
ggplot(rpriors) + geom_histogram(aes(value)) + facet_wrap(~variable, scale="free")
````

Define as curves

```{r}
s2_prior <- function(x) dinvgamma(x, s2.p[1], s2.p[2])
tau2_prior <- function(x) dinvgamma(x, tau2.p[1], tau2.p[2])
d_prior <- function(x) dgamma(x, d.p[1], scale = d.p[2]) + dgamma(x, d.p[3], scale = d.p[4])
nug_prior <- function(x) dgamma(x, nug.p[1], scale = nug.p[2]) + dgamma(x, nug.p[3], scale = nug.p[4])
beta0_prior <- function(x, tau) dnorm(x, 0, tau)

xx <- seq(.0001, 5, length.out=100)
priors <- data.frame(x = xx, s2 = s2_prior(xx), tau2 = tau2_prior(xx), beta0 = beta0_prior(xx, 1), nug = nug_prior(xx), d = d_prior(xx))
priors <- melt(priors, id="x")
ggplot(priors) + geom_line(aes(x, value)) + facet_wrap(~variable, scale="free")
```

```{r}
gp <- bgp(X=x, XX=X, Z=y, verb=0,
          meanfn="constant", bprior="b0", BTE=c(1,2000,1), m0r1=FALSE, 
          corr="exp", trace=TRUE, beta = 0,
          s2.p = s2.p, d.p = d.p, nug.p = nug.p,
          s2.lam = "fixed", d.lam = "fixed", nug.lam = "fixed", 
          tau2.lam = "fixed", tau2.p = tau2.p)
```

@@

```{r}
V <- gp$ZZ.ks2
tgp_dat <- data.frame(x   = gp$XX[[1]], 
                  y   = gp$ZZ.km, 
                 ymin = gp$ZZ.km - 1.96 * sqrt(gp$ZZ.ks2), 
                 ymax = gp$ZZ.km + 1.96 * sqrt(gp$ZZ.ks2))
```

@@

```{r gp-plot}
ggplot(tgp_dat) +
    geom_ribbon(aes(x, y, ymin = ymin, ymax = ymax), fill="red", alpha = .1) + # Var
    geom_line(aes(x, y), col="red") + # mean
    geom_point(data = data.frame(x = x, y = y), aes(x, y)) + # raw data
    geom_ribbon(dat = manual_dat, aes(x, y, ymin = ymin, ymax = ymax), fill = "blue", alpha = .1) + # Var
    geom_line(dat = manual_dat, aes(x, y), col = "blue")  + #MEAN    
    theme_bw() + theme(plot.background = element_rect(fill = "transparent",colour = NA))
```

Look at the actual parameter estimates from the trace

```{r}
gp$trace$XX[[1]][1999,]
```

Posteriors

```{r}
require(reshape2)
hyperparameters <- c("index", "s2", "tau2", "beta0", "nug", "d", "ldetK")
posteriors <- melt(gp$trace$XX[[1]][,hyperparameters], id="index")
ggplot(posteriors) + geom_histogram(aes(value)) + facet_wrap(~variable, scales="free")

## Compare to priors with `scale` misplaced as `rate` in `d`, `nug`

Define as random samples (avoids having to specify range)

```{r}
s2_rprior <- function(x) rinvgamma(x, s2.p[1], s2.p[2])
tau2_rprior <- function(x) rinvgamma(x, tau2.p[1], tau2.p[2])
d_rprior <- function(x) rgamma(x, d.p[1], d.p[2]) + rgamma(x, d.p[3], d.p[4])
nug_rprior <- function(x) rgamma(x, nug.p[1], nug.p[2]) + rgamma(x, nug.p[3], nug.p[4])
beta0_rprior <- function(x, tau) rnorm(x, 0, tau)
n = 1000
rpriors <- data.frame(s2 = s2_rprior(n), tau2 = tau2_rprior(n), beta0 = beta0_rprior(n, 1), nug = nug_rprior(n), d = d_rprior(n))
rpriors <- melt(rpriors)
ggplot(rpriors) + geom_histogram(aes(value)) + facet_wrap(~variable, scale="free")
````

```{r}
s2_prior <- function(x) dinvgamma(x, s2.p[1], s2.p[2])
tau2_prior <- function(x) dinvgamma(x, tau2.p[1], tau2.p[2])
d_prior <- function(x) dgamma(x, d.p[1], d.p[2]) + dgamma(x, d.p[3], d.p[4])
nug_prior <- function(x) dgamma(x, nug.p[1], nug.p[2]) + dgamma(x, nug.p[3], nug.p[4])
beta0_prior <- function(x, tau) dnorm(x, 0, tau)

xx <- seq(.0001, 2500, length.out=1000)
priors <- data.frame(x = xx, s2 = s2_prior(xx), tau2 = tau2_prior(xx), beta0 = beta0_prior(xx, 1), nug = nug_prior(xx), d = d_prior(xx))
priors <- melt(priors, id="x")
ggplot(priors) + geom_line(aes(x, value)) + facet_wrap(~variable, scale="free")
```
